Patient movements

### 2024-01-19 analysis/clinical-data.Rmd
revised model and new simulation function

modeled_days <- list.files("../data-clean/patient-tracking/linked-and-imputed")
modeled_days <- gsub("[[:punct:]]rds", "", modeled_days)
modeled_days <- as.Date(modeled_days)
```

@@

### Diagnosed TB over time

```{r}
diag_tb <- df %>%
  mutate(`Date modeled` = ifelse(date %in% modeled_days, "Yes", "No")) %>%
  group_by(date, `Date modeled`) %>%
  summarize(n_infectious = sum(is_infectious)) %>%
  ungroup()

diag_tb %>%
  ggplot(aes(x = date, y = n_infectious, fill = `Date modeled`)) +
  geom_bar(stat = "identity") +
  scale_x_date(date_labels = "%b %d", breaks = modeled_days) +
  scale_y_continuous(expand = c(0,0)) +
  theme_custom() +
  theme(legend.position = "top")
```

@@

#' Assumption:
#' Number of diagnosed ~ number of diagnosed TB
#' --> Assume a Dirichlet prior with support according to the counts in the range (1,5).
```

### Suspected TB over time

```{r}
df %>%
  mutate(`Date modeled` = ifelse(date %in% modeled_days, "Yes", "No")) %>%
  group_by(date, `Date modeled`) %>%
  summarize(n_infectious = sum(was_suspected & !is_infectious)) %>%
  ungroup() %>%
  ggplot(aes(x = date, y = n_infectious, fill = `Date modeled`)) +
  geom_bar(stat = "identity") +
  scale_x_date(date_labels = "%b %d", breaks = modeled_days) +
  scale_y_continuous(expand = c(0,0)) +
  theme_custom() +
  theme(legend.position = "top")
```

>>>>

modeled_days <- list.files("../data-clean/patient-tracking/linked-and-imputed")
modeled_days <- gsub("[[:punct:]]rds", "", modeled_days)
modeled_days <- as.Date(modeled_days)
modeled_days_abbrv <- format(modeled_days, format = "%b %d")
```

@@

### Diagnosed/Suspected TB over time

```{r}
diag_tb <- df %>%
  mutate(date_in_model = format(as.Date(date), format = "%b %d"),
         date_in_model = ifelse(date_in_model %in% modeled_days_abbrv, date_in_model, "Not modeled"),
         date_in_model = factor(date_in_model, levels = c(modeled_days_abbrv, "Not modeled"))) %>%
  group_by(date, date_in_model) %>%
  summarize(n_infectious = sum(is_infectious),
            n_suspected = sum(was_suspected & !is_infectious)) %>%
  ungroup() %>%
  reshape2::melt(c("date", "date_in_model")) %>%
  mutate(variable = ifelse(grepl("infectious", variable), "Diagnosed", "Suspected"),
         variable = factor(variable, levels = c("Suspected", "Diagnosed")))

mean_diag <- round(median(diag_tb$value[diag_tb$variable=="Diagnosed"]), 2)
lower_diag <- round(quantile(diag_tb$value[diag_tb$variable=="Diagnosed"], .25), 2)
upper_diag <- round(quantile(diag_tb$value[diag_tb$variable=="Diagnosed"], .75), 2)
annon_diag <- paste0("Median diagnosed = ", mean_diag, " (IQR ", lower_diag, " – ", upper_diag, ")")

diag_tb_pl <- diag_tb %>%
  ggplot(aes(x = date, y = value, fill = date_in_model, alpha = variable)) +
  geom_bar(position = position_stack(), stat = "identity") +
  annotate("text", x = as.Date("2021-11-04"), y = 12, label = annon_diag, size = 14/cm(1)) +
  scale_x_date(date_labels = "%b %d", breaks = modeled_days) +
  scale_y_continuous(expand = c(0,0)) +
  scale_fill_manual(values = c(wes_palette("Moonrise3"), "grey")) +
  scale_alpha_discrete(breaks = c("Diagnosed", "Suspected"), range = c(.33, 1)) +
  labs(y = "Number of patients",
       title = "Daily number of TB patients attending the clinic",
       subtitle = "Diagnosed and suspected (tested) patients in October and November") +
  theme_custom(14) +
  theme(legend.position = "bottom",
        legend.title = element_blank(),
        legend.direction = "vertical", 
        legend.box = "vertical",
        axis.title.x = element_blank(),
        axis.text.x = element_blank()) +
  guides(fill = guide_legend(nrow = 1, order = 2),
         alpha = guide_legend(nrow = 1, order = 1))

save_plot(diag_tb_pl, pdf_file = "../results/data/diagnosed-tb-patients.pdf", w = 18, h = 12)
```

@@

#' Assumption:
#' Number of diagnosed ~ number of diagnosed TB
#' --> Assume a Dirichlet prior with support according to the counts in the range (1,5).

table(diag_tb$n_infectious)

source("../utils/trans_risk.r")

undiag_tb_pl <- data.frame(x = rundiag(1e5)) %>%
  group_by(x) %>%
  summarize(n = n()) %>%
  ungroup() %>%
  mutate(p = 100 * n / sum(n),
         p = round(p)) %>%
  ggplot(aes(x = factor(x), y = p)) +
  geom_bar(stat = "identity", color = wes_palette("GrandBudapest1")[3], fill = wes_palette("GrandBudapest1")[4], alpha = .5) +
  labs(x = "Number of undiagnosed TB patients", y = "Probability (%)",
       title = "Undiagnosed TB patients among clinical attendees",
       subtitle = "Assuming that no. undiagnosed similar to no. diagnosed",
       caption = "Note: Distribution derived from empirical clinical data") +
  coord_cartesian(expand = F) +
  scale_y_continuous(limits = c(0,NA)) +
  theme_custom(14) +
  theme(plot.caption = element_text(size = 10))

undiag_tb_pl
save_plot(undiag_tb_pl, pdf_file = "../results/data/undiagnosed-tb-patients.pdf", w = 14, h = 12)
```

### 2024-01-19 analysis/environmental-data.Rmd
revised model and new simulation function

```{r}
dates <- sort(unique(cont_df$date))
abbr_dates <- paste(month(dates, label = T, abbr = T), day(dates))
cont_df <- readRDS("../data-clean/environmental/continuous-iaq-data.rds") %>%
  mutate(time = as.POSIXct(format(date_time, format = "%H:%M:%S"), format = "%H:%M:%S"),
         date = factor(date, levels = dates))
aer <- readRDS("../data-clean/environmental/air-change-rate.rds") %>%
  mutate(mth = month(date, label = T, abbr= T),
         dy = day(date),
         mthdy = paste(mth, dy),
         mthdy = factor(as.character(mthdy), levels = abbr_dates),
         daytime = factor(daytime, levels = c("Morning", "Afternoon")))
```

@@

```{r}
cont_df %>%
  ggplot(aes(x = time, y = co2, color = date)) +
  geom_step() +
  facet_wrap(~ room) +
  coord_cartesian(expand = F) +
  scale_x_datetime(date_breaks = "1 hour", date_labels = "%H:%M") +
  scale_color_manual(values = wes_palette("Moonrise3")) +
  labs(y = "Number of people in waiting room") +
  theme_custom() +
  theme(axis.title.x = element_blank(),
        legend.title = element_blank())
```

@@

```{r}
aer %>%
  ggplot(aes(x = mthdy, fill = daytime, y = aer)) +
  geom_bar(stat = "identity", position = position_dodge(width = 1, preserve = "single"), width = .66) +
  facet_wrap(~ room) +
  scale_fill_brewer(palette = 9) +
  scale_y_continuous(expand = expansion(mult = c(0, .05))) +
  theme_custom()
```

>>>>

```{r}
dates <- as.Date(c("2021-10-13", "2021-10-15", "2021-10-25", "2021-11-04", "2021-11-05"))
abbr_dates <- format(dates, format = "%b %d")
cont_df <- readRDS("../data-clean/environmental/continuous-iaq-data.rds") %>%
  mutate(time = as.POSIXct(format(date_time, format = "%H:%M:%S"), format = "%H:%M:%S"),
         mthdy = factor(as.character(format(date, format = "%b %d")), levels = abbr_dates))
aer <- readRDS("../data-clean/environmental/air-change-rate.rds") %>%
  mutate(mthdy = factor(as.character(format(date, format = "%b %d")), levels = abbr_dates),
         daytime = factor(daytime, levels = c("Morning", "Afternoon")))
```

@@

```{r}
co2_pl <- cont_df %>%
  ggplot(aes(x = time, y = co2, color = mthdy)) +
  geom_step() +
  facet_wrap(~ room) +
  coord_cartesian(expand = F) +
  scale_x_datetime(date_breaks = "3 hours", date_labels = "%H:%M") +
  scale_color_manual(values = wes_palette("Moonrise3")) +
  labs(y = expression("CO"[2]*" (parts per million, ppm)"),
       title = "Carbon dioxide measurements",
       subtitle = expression("CO"[2]*" levels over time per room")) +
  theme_custom(14) +
  theme(axis.title.x = element_blank(),
        legend.title = element_blank(),
        legend.position = "bottom")

save_plot(co2_pl, pdf_file = "../results/data/co2-levels-over-time.pdf", w = 18, h = 12)
```

@@

```{r}
aer_pl <- aer %>%
  ggplot(aes(x = room, fill = daytime, y = aer)) +
  geom_boxplot(position = position_dodge(width = .85)) +
  #geom_bar(stat = "identity", position = position_dodge(width = 1, preserve = "single"), width = .66) +
  #facet_wrap(~ room) +
  scale_fill_brewer(palette = 9) +
  scale_y_continuous(expand = expansion(mult = c(0, .05)), limits = c(0, NA)) +
  labs(y = "Air change rate (changes per hour)",
       title = "Ventilation conditions in the clinic",
       subtitle = expression("Air exchange rate as computed from CO"[2])) +
  theme_custom(14) +
  theme(axis.title.x = element_blank(),
        legend.title = element_blank(),
        legend.position = c(0.8, 0.9))

save_plot(aer_pl, pdf_file = "../results/data/air-changes-per-hour.pdf", w = 11, h = 12)
```

### 2024-01-19 analysis/patient-tracking.Rmd
revised model and new simulation function

source("../utils/plotting.r")
source("../utils/spatial.r")

text_descr <- 8
```

@@

```{r}
source("../helper/building.R")
```

@@

```{r}
df %>%
  group_by(date, patient_id) %>%
  slice(1) %>%
  ungroup() %>%
  dplyr::select(tracking_end) %>%
  unlist() %>%
  table()

@@

time_in_clin %>%
  ggplot(aes(x = tracking_end, y = total_duration)) +
  geom_boxplot()

time_in_clin %>%
  filter(tracking_end != "Possible HCW") %>%
  dplyr::select(-tracking_end) %>%
  melt(c("date", "patient_id")) %>%
  mutate(variable = factor(variable, levels = c("Clinic", "Waiting room", "TB room", "Corridor"))) %>%
  ggplot(aes(x = variable, fill = variable, y = value)) +
  geom_boxplot() +
  scale_y_sqrt()
```

@@

close_contact %>%
  filter(tracking_end != "Possible HCW") %>%
  ggplot(aes(x = Clinic, y = n_cc)) +
  geom_point()

close_contact %>%
  filter(tracking_end != "Possible HCW") %>%
  ggplot(aes(x = 1, y = cc_time / Clinic)) +
  geom_boxplot() 
```

@@

```{r}
clinic_occupancy <- df %>%
  group_by(date, time) %>%
  summarize(
    `Clinic` = n(),
    `Waiting room` = sum(is_waitingroom),
    `TB room` = sum(is_tbroom),
    `Corridor` = sum(is_passage)) %>%
  ungroup() %>%
  melt(c("date", "time")) %>%
  mutate(variable = factor(variable, levels = c("Clinic", "Waiting room", "TB room", "Corridor")),
         date = format(date, "%b %d"),
         time = as.POSIXct(format(time, format = "%H:%M:%S"), format = "%H:%M:%S")) %>%
  filter(between(hour(time), 7, 16))

clinic_occupancy %>%
  ggplot(aes(x = time, y = value, color = date)) +
  geom_step() +
  facet_wrap(~ variable) +
  coord_cartesian(expand = F) +
  scale_x_datetime(date_breaks = "1 hour", date_labels = "%H:%M") +
  scale_color_manual(values = wes_palette("Moonrise3")) +
  labs(y = "Number of people in waiting room") +
  theme_custom() +
  theme(axis.title.x = element_blank(),
        legend.title = element_blank())
```

@@

```{r}
col <- RColorBrewer::brewer.pal(n = 9, name = "YlOrRd")

plot_spatial <- function(pl) {
  pl +
  ggpubr::background_image(clinic_img) +
  scale_y_continuous(expand = expansion(add=c(1700, 3000)), limits = c(-6500, 2000)) +
  scale_x_continuous(expand = expansion(add=c(400, 500)), limits = c(-16000, 7500)) +
  annotation_custom(entrance_lab, xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = Inf) +
  annotation_custom(wr_lab, xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = Inf) +
  annotation_custom(tb_lab, xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = Inf) +
  annotation_custom(re_lab, xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = Inf) +
  theme_custom() +
  theme(text = element_text(size = text_descr),
        axis.text = element_blank(), axis.title = element_blank(), axis.ticks = element_blank(),
        legend.position = c(0.825,0.05), legend.direction = "horizontal", 
        legend.key.width = unit(0.3, "cm"), legend.key.height = unit(0.2, "cm"),
        panel.background = element_rect(fill = col[1]), panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()) +
  guides(fill = guide_legend(title.position = "top"))
}

plot_spatial(
  ggplot(data = df, mapping = aes(x = x, y = y)) +
    geom_bin2d(binwidth = 200) ) +
  labs(fill = "Number of tracks") +
  scale_fill_stepsn(colours = RColorBrewer::brewer.pal(n = 9, name = "YlOrRd"), trans = "log10", 
                    breaks = c(1, 10, 100, 1000, 10000)) 
```

@@

# manual sample of tracks
ex_pid_man <- c(
  "2021-10-25 4649", # Entered + Exited, including TB room stay
  "2021-10-13 6297", # Lost Exit, only waiting room
  "2021-11-04 12860", # (Entered) + Exited, waiting room and corridor
  "2021-10-13 9346", # Entered + Exited, including TB room
  "2021-11-05 16625", # Entered + (Exited), only waiting room
  "2021-10-25 9526", # Lost Exit, waiting room and corridor
  "2021-11-04 9841", # Entered = (Exited), waiting room and corridor
  "2021-10-15 8644", # Entered + Exited, including TB room
  "2021-10-15 9419" # Lost Enter
)
ex_df_man <- df %>% filter(day_pid %in% ex_pid_man)
ex_df_man_se <- ex_df_man %>%
  group_by(day_pid) %>%
  slice(1, n()) %>%
  mutate(type = factor(c("Start", "End"), levels=c("Start", "End"))) %>%
  ungroup()

# plot
plot_spatial(
  ggplot() +
    geom_path(data = ex_df_man, mapping = aes(x = x, y = y), color = "darkred") +
    geom_point(data = ex_df_man_se, mapping = aes(x = x, y = y, shape = type), size = 2) +
    facet_wrap(~ day_pid) +
    scale_shape_manual(values = c(1,13)) 
) +
  theme(legend.position = "none")
```

>>>>

source("../utils/plotting.r")
source("../utils/spatial.r")
```

@@

```{r}
cellSize = 250
source("../preprocessing/prep_building-rasterize.R")
```

@@

```{r}
track_qual <- df %>%
  group_by(date, patient_id) %>%
  slice(1) %>%
  ungroup() %>%
  dplyr::select(tracking_end) %>%
  unlist() %>%
  table()

sum(track_qual[names(track_qual) != "Possible HCW"])
track_qual_str <- paste0(names(track_qual), ": ", track_qual, " (", round(track_qual / sum(track_qual) * 100), "%)")
track_qual_str

@@

time_in_clin_pl <- time_in_clin %>%
  filter(tracking_end != "Possible HCW") %>%
  dplyr::select(-tracking_end) %>%
  melt(c("date", "patient_id")) %>%
  mutate(variable = factor(variable, levels = c("Clinic", "Waiting room", "TB room", "Corridor"))) %>%
  ggplot(aes(x = variable, fill = variable, y = value)) +
  geom_boxplot() +
  scale_y_sqrt() +
  scale_fill_manual(values = wes_palette("Royal2")) +
  labs(y = "Time (minutes)", title = "Time spent in clinic", subtitle = "Duration of tracked clinical visit") +
  theme_custom(14) +
  theme(legend.position = "none",
        legend.title = element_blank(),
        axis.title.x = element_blank(),
        plot.title.position = "plot")

save_plot(time_in_clin_pl, pdf_file = "../results/data/time-in-clinic.pdf", w = 13, h = 12)
```

@@

cc_n_pl <- close_contact %>%
  filter(tracking_end != "Possible HCW") %>%
  ggplot(aes(x = n_cc))  +
  scale_x_continuous(expand = c(0,0)) +
  scale_y_continuous(expand = c(0,0)) +
  labs(y = "Density", x = "Number of close contacts",
       title = "Close contacts",
       subtitle = "Defined as <1m for >1min") +
  geom_density(fill = wes_palette("FantasticFox1")[2], color = wes_palette("FantasticFox1")[2], alpha = 0.8) +
  theme_custom(14) +
  theme(plot.title.position = "plot",
        plot.margin = margin(r = 11),
        plot.subtitle = element_text(margin = margin(b = 11)))
cc_n_pl

cc_t_pl <- close_contact %>%
  filter(tracking_end != "Possible HCW") %>%
  ggplot(aes(x = cc_time / Clinic * 100) ) +
  scale_x_continuous(expand = c(0,0), limits = c(0, 100)) +
  scale_y_continuous(expand = c(0,0)) +
  labs(y = "Density", x = "Time in close contact (%)", title = "", subtitle = "") +
  geom_density(fill = wes_palette("Moonrise1")[2], color = wes_palette("Moonrise1")[2], alpha = 0.8) +
  theme_custom(14) +
  theme(plot.title.position = "plot",
        plot.margin = margin(r = 11))
cc_t_pl

cc_pl <- (cc_n_pl + cc_t_pl) + plot_layout(nrow = 1, heights = c(0.5, .5)) & 
  plot_annotation(theme = theme(plot.background = element_rect(fill = "transparent", color = "transparent")))
save_plot(cc_pl, pdf_file = "../results/data/close-contact.pdf", w = 16, h = 12)
```

@@

```{r}
clinic_occupancy <- df %>%
  group_by(date, time) %>%
  summarize(
    `Clinic` = n(),
    `Waiting room` = sum(is_waitingroom),
    `TB room` = sum(is_tbroom),
    `Corridor` = sum(is_passage)) %>%
  ungroup() %>%
  melt(c("date", "time")) %>%
  mutate(variable = factor(variable, levels = c("Clinic", "Waiting room", "TB room", "Corridor")),
         date = format(date, "%b %d"),
         date = factor(as.character(date), levels = c("Oct 13", "Oct 15", "Oct 25", "Nov 04", "Nov 05")),
         time = as.POSIXct(format(time, format = "%H:%M:%S"), format = "%H:%M:%S")) %>%
  filter(between(hour(time), 7, 16))

time_pl <- clinic_occupancy %>%
  ggplot(aes(x = time, y = value, color = date)) +
  geom_step() +
  facet_wrap(~ variable) +
  coord_cartesian(expand = F) +
  scale_x_datetime(date_breaks = "3 hours", date_labels = "%H:%M") +
  scale_color_manual(values = wes_palette("Moonrise3")) +
  labs(y = "Number of people", title = "Occupancy over time", subtitle = "Number of active tracks per room over time") +
  theme_custom(14) +
  theme(axis.title.x = element_blank(),
        legend.title = element_blank(),
        legend.position = "bottom"
        )

save_plot(time_pl, pdf_file = "../results/data/no-people-over-time.pdf", w = 16, h = 12)
```

@@

```{r}
spatial_pl <- plot_spatial(
  ggplot(data = df, mapping = aes(x = x, y = y)) +
    geom_bin2d(binwidth = 200),
  text_descr = 14) +
  labs(fill = "Number of tracks", title = "Spatial occupancy", subtitle = "Total number of tracks per grid cell") +
  scale_fill_stepsn(colours = RColorBrewer::brewer.pal(n = 9, name = "YlOrRd"), trans = "log10", 
                    breaks = c(1, 10, 100, 1000, 10000)) +
  theme(legend.position = c(0.825,0.075), legend.direction = "horizontal") +
  guides(fill = guide_legend(nrow = 2, title.position = "top"))

save_plot(spatial_pl, pdf_file = "../results/data/no-people-spatial.pdf", w = 16, h = 12)
```

@@

# manual sample of tracks
ex_pid_man <- c(
  "2021-10-25 4649", # Entered + Exited, including TB room stay
  "2021-10-13 6297", # Lost Exit, only waiting room
  "2021-11-04 12860", # (Entered) + Exited, waiting room and corridor
  "2021-10-13 9346", # Entered + Exited, including TB room
  "2021-11-05 16625", # Entered + (Exited), only waiting room
  #"2021-10-25 9526", # Lost Exit, waiting room and corridor
  #"2021-11-04 9841", # Entered = (Exited), waiting room and corridor
  #"2021-10-15 8644", # Entered + Exited, including TB room
  "2021-10-15 9419" # Lost Enter
)
ex_df_man <- df %>% filter(day_pid %in% ex_pid_man)
ex_df_man_se <- ex_df_man %>%
  group_by(day_pid) %>%
  slice(1, n()) %>%
  mutate(type = factor(c("Start", "End"), levels=c("Start", "End"))) %>%
  ungroup()

# plot
ex_tracks <- plot_spatial(
  ggplot() +
    geom_path(data = ex_df_man, mapping = aes(x = x, y = y), color = "darkred") +
    geom_point(data = ex_df_man_se, mapping = aes(x = x, y = y, shape = type), size = 2) +
    facet_wrap(~ day_pid, ncol = 3) +
    scale_shape_manual(values = c(1,13)),
  text_descr = 10
) +
  labs(title = "Sample of invidiual tracks", subtitle = "Illustrative examples of the mixed tracking quality") +
  theme(legend.position = "none",
        plot.title = element_text(size = 16),
        plot.subtitle = element_text(size = 14))

save_plot(ex_tracks, pdf_file = "../results/data/example-patient-tracks.pdf", w = 22, h = 12)
```
